
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##A day with ggplot2 and playing with R Markdown

```{r iris data}
#Load library
library(data.table)
library(ggplot2)
library(gridExtra)
DT <- data.table(iris)
```

```{r attaching data,echo=FALSE}
attach(DT)
```

##Data Exploration
There are `r nrow(DT)` specimens of flowers belonging to `r levels(Species)`
```{r iris EDA}
head(DT)
summary(DT)

```


## Including Plots

You can also embed plots, for example:
```{r qplot,echo=FALSE}
qplot(x = Sepal.Length, y = Sepal.Width , data = DT, color = Species)
```

# Adding smooth geom
```{r smooth geom, echo=FALSE}
qplot(x = Sepal.Length, y = Sepal.Width , data = DT, geom = c("point","smooth"))
```

#Histogram of sepal length frequency 
```{r histogram, echo=FALSE , results="hide"}
qplot(Sepal.Length, data = DT, fill = Species)
```


#Scartterplots with facets
```{r scatterplots with facets, echo=FALSE}
par(mfrow=c(2,2))
plot1 <- qplot(Sepal.Length,Sepal.Width, data = DT, facets = .~ Species,color = Species)
plot2 <- qplot(Petal.Length,Petal.Width, data = DT, facets = .~ Species,color = Species)
grid.arrange(plot1, plot2)

```


##Lets move on to 'The ggplot2'
I chose ggplot2 because of its use of grammar of graphics and high level of flexiblity in plot specification.  

###Grammar of graphics
The basic idea of grammar of graphics is to independently specify plot building blocks and finally combine the blocks to get the desired graphs and plots you want. Also, the 
Building blocks include:

* data
* aesthetic mapping
* geometric object
+ statistical transformations
* scales
* coordinate system
* position adjustments
* faceting

####Let's get to the data
```{r}
str(DT)
head(DT)
```

####Basic Aesthetic Mapping:
```{r}
p <- ggplot(DT, aes(x= hp, y= mpg))
```

####Adding a geometic object( In this case, points)
```{r}
p + geom_point()
```

####Smoothers
```{r}
ggplot(DT, aes(mpg)) + geom_histogram(stat="bin", binwidth = 1)

p + geom_point() +geom_smooth(method = "lm")
p + geom_point() +geom_smooth(method = "auto")
p + geom_point() +geom_smooth(span = .4)
```

#### Scales- For controlling aesthetic mapping
```{r}
p4 <- p3 +geom_point(aes(color = am),
                     alpha = 0.5,
                     size = 1.5,
                     position = position_jitter(width = 0.25, height = 0))
```
```{r}
p4 + scale_color_continuous(name = "Auto/Manual",
                         breaks = c(0.0,1.0),
                         labels = c(0,1))
```

####Change color to low and high value

```{r}

p4 + scale_color_continuous(name = "Auto/Manual",
                            breaks = c(0.0,1.0),
                            labels = c(0,1),
                            low = "red",
                            high = "green")
```



